Growing Up with Nell: A Narrative Interface for Literacy

### 2012-03-12 paper.tex
Massage the 'Extensible dialog' section.

\subsection{Extensible dialog}
The child can directly interact with the Princess and other agents within
Nell.  Convincing discourse is a hard problem, but
reasonable approximations do not have to be difficult; even
the rudimentary conversational abilities of ELIZA elicited hours of
conversation~\cite{weizenbaum:power}.
% her secretary is divulging her life story
% asks him to leave the room so he could be alone with it
% hours discussing her boyfriend.
Nell's use as an interactive diary is enhanced by the child's
collaboration in the fiction that Nell is intelligent---what
Turkle~\cite{turkle:alone} calls the ``ELIZA effect.''

@@

We leverage AIML~\cite{aiml:2005},
the markup language developed for the three-time Loebner Prize-winning
AliceBot.%
\footnote{Wilcox identifies several shortcomings with the expressive
  power of AIML, proposing ChatScript as a
  replacement~\cite{wilcox:2010}.  AIML's community and
  diversity of third-party implementations make it preferable at present.}
Each story module can include AIML fragments extending the
conversational capabilities of Nell's agents.

Fig.~\ref{fig:aiml}
shows how this can be used to implement a traditional
text adventure interaction for a particular story module.
In this example the child can write instructions such as, ``go north,'' ``go
east,'' or ``open door'' for the agent.  These trigger events
handled by the story module to move the agent through the
adventure---for example, by causing the activity to draw a
new location, causing the agent to describe it, and shifting the
AIML topic to enable vocabulary suitable for the new location.

>>>>

\subsection{Extensible dialog}
Nell's use as an interactive diary and
Portfolio~\cite{stefanakis:portfolios} is enhanced by the user's
collaboration in the fiction that Nell is intelligent---what
Turkle~\cite{turkle:alone} calls the ``ELIZA effect.''
Convincing discourse is a hard problem, but
reasonable approximations do not have to be difficult; even
the rudimentary conversational abilities of ELIZA elicited hours of
conversation~\cite{weizenbaum:power}.
% her secretary is divulging her life story
% asks him to leave the room so he could be alone with it
% hours discussing her boyfriend.
We've chosen to allow the child to directly converse with the
Princess and other agents within Nell.

@@

Our implementation extends AIML~\cite{aiml:2005},
the markup language developed for the three-time Loebner Prize--winning
AliceBot.%
\footnote{Wilcox identifies several shortcomings with the expressive
  power of AIML, proposing ChatScript as a
  replacement~\cite{wilcox:2010}.  AIML's community and
  diversity of third-party implementations make it our choice at present.}
Each story module can include AIML fragments extending the
conversational capabilities of Nell's agents.

Figure~\ref{fig:aiml}
shows an AIML fragment augmenting an agent with
commands suitable for a text adventure story module.
In such a story the child might write instructions such as
\texttt{go north}, \texttt{go east}, or \texttt{open door} in
addition to usual agent dialog.
These new commands fire events which are referenced by scene
preconditions.  For example, \texttt{north} may
trigger a scene which causes the activity to draw a new location on
the page, causes the agent to write and speak a description of the new
location, and which shifts the AIML topic to enable additional vocabulary
suitable for the new location.
